Stochastic Variance-Reduced Policy Gradient

### 2018-02-05 svrpg.tex
Updated
Updated algorithms and notation

Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^d\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$. We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.
The search for a locally optimal policy is performed by means of gradient ascent, where the policy gradient
%of $J(\vtheta)$ \wrt the policy parameters 
is \cite{sutton2000policy, Peters2008reinf}:
\begin{align} \label{E:policygradient}
        \gradJ{\vtheta} = \EVV[\tau \sim \pol]{\score{\vtheta}{\tau}\Reward(\tau)}
        %\int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
\end{align}
Notice that the distribution defining the gradient is induced by the current policy. This aspect introduces a nonstationarity in the sampling process. Since the underlying distribution changes over time, it is necessary to resample at each update or use weighting techniques such as importance sampling.
Here, we consider the \emph{online learning scenario}, where trajectories are sampled by interacting with the environment at each policy change. 
In this setting, stochastic gradient ascent is typically employed.
At each iteration $k >0$, a batch $\mathcal{D}_N^k = \{\tau_i\}_{i=0}^N$ of $N>0$ trajectories is collected using policy $\pi_{\vtheta_k}$.
The policy is then updated as $\vtheta_{k+1}  = \vtheta_k + \alpha\gradApp{\vtheta_k}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of~\eqref{E:policygradient} using $\mathcal{D}_N^k$, \ie
\begin{align} \label{E:policygradient.estimate}
        \gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N} g(\tau_i|\vtheta), \quad \tau_i \in \mathcal{D}_N^k.
\end{align}
The most widespread gradient estimators are REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
REINFORCE is such that $g(\tau|\vtheta) = \nabla \log\pi(\tau|\vtheta) \mathcal{R}(\tau)$,  we refer the reader to App.~\ref{A:gradient_estimators} for details.

@@

The idea of SVRG is to alternate full and stochastic gradient updates. 
Each $m = \mathcal{O}(N)$ iterations, a snapshot $\widetilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\widetilde{\vtheta}) = \frac{1}{N} \sum_i z(x_i|\widetilde{\vtheta})$.
Between snapshots, the parameter is updated with a gradient estimate corrected using stochastic gradient, for any $t \in \{0,\ldots,m-1\}$
\begin{equation}\label{E:svrg.gradient.correction}
        \blacktriangledown f(\vtheta_{t+1}) = \nabla f(\tilde{\vtheta}) + \nabla z(x_i | \vtheta_t) - \nabla z(x_i | \tilde{\vtheta}),
\end{equation} 
where $x_i$ is sampled uniformly at random from $\mathcal{D}_N$ (\ie $x_i \sim \mathcal{U}(\mathcal{D}_N)$). The corrected gradient $\blacktriangledown f(\vtheta)$ is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, without resorting to plain full gradient. The resulting IFO complexity is $\mathcal{O}(N+\nicefrac{\sqrt{N}}{\epsilon})$ \cite{reddi2016stochastic}.

More recently, some extensions of variance reduction algorithms to the non-convex objectives have been proposed~\citep[\eg][]{reddi2016stochastic,allen2016variance,reddi2016stochastic,reddi2016fast}. In this scenario, $f$ is typically required to be L-smooth, \ie $\norm[]{\nabla f(\vtheta') - \nabla f(\vtheta)} \leq L\norm[]{\vtheta'-\vtheta}$ for each $\vtheta,\vtheta'\in\realspace^n$ and for some Lipschitz constant $L$. Under this hypothesis, the IFO complexity of full gradient  and of stochastic gradient ascent are the same as in the \textcolor{red}{concave} case~\citep{nesterov2013introductory,ghadimi2013stochastic}. Also in this case, SVRG yields an advantageous trade-off, with an IFO complexity of $\mathcal{O}(N + \nicefrac{N^{\nicefrac{2}{3}}}{N})$~\citep{reddi2016stochastic}. The only additional requirement is to select $\vtheta^*$ uniformly at random among all the $\vtheta_k$ instead of simply setting it to the final value. Pseudocode of SVRG for both the convex and the non-convex case is provided in Alg.~\ref{alg:svrg}. 
\todopir{SVRG for non-convex optimization requires to mix the parameter $\wt{\vtheta}^{s+1}$?}

\begin{algorithm}[tb]
	\caption{SVRG}
	\label{alg:svrg}
	\begin{algorithmic}
            \STATE {\bfseries Input:} a dataset $\mathcal{D}_N$, number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m}^0$
		\FOR{$s=0$ {\bfseries to} $S-1$}
		\STATE $\vtheta_0^{s+1} := \wt{\vtheta}^s := \vtheta_{m}^s$
		\FOR{$t=0$ {\bfseries to} $m-1$}
%		\STATE Pick $i$ uniformly at random from $[1,N]$
        \STATE $x_i \sim \mathcal{U}\left(\mathcal{D}_N\right)$
		\STATE $v^{s+1}_t = 
			\nabla f(\wt{\vtheta}^s) + 
			\nabla g(x_i|\vtheta_t^{s+1}) -
			\nabla g(x_i|\wt{\vtheta}^{s})
		$
        \STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}_t$
		\ENDFOR
		\ENDFOR
		\STATE {\bfseries Convex case: return} $\vtheta_{m}^S$
		\STATE {\bfseries Non-Convex case: return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $[0,S)\times[0,m)$
	\end{algorithmic}
\end{algorithm}

@@

We can now introduce Stochastic Variance-Reduced Policy Gradient (SVRPG).
% For simplicity, we focus on the REINFORCE estimator but the extension to G(PO)MDP is straightforward.
Between snapshots, the SVRPG estimate is corrected using importance weighting, for any $t \in \{0,\ldots,m-1\}$
\begin{align*}
        \blacktriangledown J(\vtheta_{t+1}) &= \wh{\nabla}_N J(\wt{\vtheta}) + g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
\end{align*}
% \begin{align*}
% 	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) \\
% 		&+ \nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
% 		- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i),
% \end{align*}
where $\wh{\nabla}_N J(\wt{\vtheta})$ is as in~\eqref{E:policygradient.estimate} where $\mathcal{D}_N$ is sampled using the snapshot policy $\pi_{\wt{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pi_{\vtheta_t}$, and $\omega(\tau|\vtheta_t, \wt{\vtheta}) = \frac{\pi(\tau|\wt{\vtheta})}{\pi(\tau|\vtheta_t)}$ is an importance weight from $\pi_{\vtheta_t}$ to the snapshot policy $\pi_{\wt{\vtheta}}$. 
Our update is still fundamentally on-policy, since the weighting concerns only the correction term. However, this partial ``off-policyness'' is enough to introduce variance. This is a well-known issue of importance sampling~\citep[\eg][]{thomas2015high}. To mitigate it, we use mini-batches of trajectories of size $B \ll N$ to average the correction, \ie
\begin{align}\label{E:svrpg.estimate.batch}
        \blacktriangledown J(\vtheta_{t+1}) &= \wh{\nabla}_N J(\wt{\vtheta})\\ \notag
                                            & \quad{} + \frac{1}{B} \sum_{i=0}^{B-1} \left[
        g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
        \right].
\end{align}
%\begin{align*}
%&\blacktriangledown J(\vtheta) = \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) +\\
%&\quad\frac{1}{B}\left[\sum_{i=0}^{B-1}
%\nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
%- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i)\right].
%\end{align*}
The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case~\citep{harikandeh2015stopwasting,konevcny2016mini}. It is easy to show that the SVRPG estimator has the following, desirable properties (see App.~\ref{app:proofs}):
\begin{lemma}
Denote by $\wh{\nabla}_N^{\textsc{RF}} J(\vtheta)$ the REINFORCE estimator in~\eqref{E:policygradient.estimate} and let $\vtheta^* \in \argmin_{\vtheta} \{J(\vtheta)\}$. Then, the SVRG estimate in~\eqref{E:svrpg.estimate.batch} is \emph{unbiased}
\[
\mathop{\mathbb{E}}
\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
\]
and regardless of the mini-batch size $B$:
\[
	\mathbb{V}ar\left[\gradBlack{\vtheta^*}\right] = 
    \mathbb{V}ar\left[\wh{\nabla}_N^{\textsc{RF}}J(\vtheta^*)\right].
\]
\todopir{Definition of Var??}
\end{lemma}

%\begin{restatable}{lemma}{unbias}\label{lemma:unbias}
%The SVRPG estimator is unbiased:
%\[
%\mathop{\mathbb{E}}
%\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
%\]
%\end{restatable}
%\begin{restatable}{lemma}{zerovar}\label{lemma:zerovar}
%At convergence, the variance of the SVRPG estimator is equal to the variance of the REINFORCE estimator, regardless of the mini-batch size $B$:
%\[
%	\mathbb{V}ar\left[\gradBlack{\vtheta^*}\right] = 
%	\mathbb{V}ar\left[\gradApp{\vtheta^*}{N}\right].
%\]
%\end{restatable}
In particular, the latter result suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can actually achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance (at least asymptotically).
The resulting SVRPG policy optimization method is detailed in Alg.~\ref{alg:svrpg}.
To obtain a method that can be used in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient, the mini-batch size $B$, and the step size $\alpha$.
Adaptive selection of these meta-parameters is also possible, and is explored in Sec.~\ref{sec:stopping}.
However, the next section will provide general converge guarantees for SVRPG as in Alg.~\ref{alg:svrpg}.

\begin{algorithm}[tb]
        \caption{SVRPG based on REINFORCE}
	\label{alg:svrpg}
	\begin{algorithmic}
		\STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{m}^0$, batch size $N$, mini-batch size B
		\FOR{$s=0$ {\bfseries to} $S-1$}
        \STATE Sample $N$ trajectories $\{\tau_j\}$ from $\pi(\cdot\vert\tilde{\vtheta}^{s})$
		\STATE $\gradApp{\tilde{\vtheta}^{s}}{N} = \frac{1}{N}\sum_{j=0}^{N-1}\score{}{\tau_j\vert\tilde{\vtheta}^{s}}\Reward(\tau_j)$
		\STATE $\vtheta_0^{s+1} = \tilde{\vtheta}^s = \vtheta_{m}^s$
		\FOR{$t=0$ {\bfseries to} $m-1$}
        \STATE Sample $B$ trajectories $\{\tau_i\}$ from $\pi(\cdot\vert\vtheta_t^{s+1})$
        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1}
        \begin{aligned}[t]
                & \Reward(\tau_i) \Big[ \score{}{\tau_i|\vtheta_t^{s+1}} \\
                &- \omega(\tau|\vtheta^{s+1}_t, \wt{\vtheta}^s)\score{}{\tau_i|\wt{\vtheta}^{s}} \Big]
        \end{aligned}$
        \STATE $v^{s+1}_t = \gradApp{\wt{\vtheta}^{s}}{N} + c^{s+1}_t$
%%		\begin{align*}
%%		\blacktriangledown J(\vtheta_t^{s+1}) = 
%%		&\gradApp{\tilde{\vtheta}^s}{N} \\
%%		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
%%		\score{}{\tau_i\vert\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
%%		&\left. - \omega(\tau_i)\score{}{\tau_i \vert \tilde{\vtheta}^{s}}\Reward(\tau_i)\right]
%%		\end{align*}
        \STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}_t$
		\ENDFOR
		\ENDFOR
		\STATE {\bfseries return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $[0,S)\times[0,m)$
	\end{algorithmic}
\end{algorithm}

\section{Convergence Guarantees of SVRPG}\label{sec:conv}
In this section, we state the convergence guarantees for SVRPG with REINFORCE structure (see Alg.~\ref{alg:svrpg}).
We mainly leverage on the recent analysis of \textcolor{red}{non-concave} SVRG~\cite{reddi2016stochastic,allen2016variance}.
Each of the three challenges presented at the beginning of Sec.~\ref{sec:alg} can potentially prevent convergence, so we need additional assumptions.
In App.~\ref{app:gauss} we show how the widespread Gaussian policy satisfies these assumptions.

@@

We can now state the convergence guarantees for SVRPG.
\todopir{Is it specific for REINFORCE?}
\begin{restatable}[Convergence of the SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
Under Ass.~\ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Alg.~\ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
\begin{align*}
	&\EVV[]
	{\norm[2]{\nabla J(\vtheta_A)}^2} 
		\leq
		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
		\frac{\zeta}{N}
		+\frac{\xi}{B}
\end{align*}
\end{restatable}
% Specific constants and meta-parameter constraints are provided in the proof of the theorem in Appendix \ref{app:proofs}.
Refer to App.~\ref{app:proofs} for a detailed proof involving definition of the constants and the meta-parameter constraints.
%The theoretical settings may be too conservative.
In Sec.~\ref{sec:stopping} we propose a joint selection of step size $\alpha$ and epoch size $m$ which is more suited for applications. 
By analysing the upper-bound in Thm.~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on \textcolor{red}{non-concave} SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
To achieve true convergence, both the batch size $N$ and the mini-batch size $B$ should increase over time. 
In practice, it is enough to choose $N$ and $B$ large enough to make the second and the third term negligible, \ie to mitigate the variance introduced by FG approximation and importance sampling, respectively.
Since the optimal configuration may be task-dependent, we leave this choice to the empirical analysis of Sec.~\ref{sec:exp}.

>>>>

Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^d\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$. We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.
The search for a locally optimal policy is performed by means of gradient ascent, where the policy gradient
%of $J(\vtheta)$ \wrt the policy parameters 
is \cite{sutton2000policy, Peters2008reinf}:
\begin{align} \label{E:policygradient}
        \gradJ{\vtheta} = \EVV[\tau \sim \pol]{\score{\vtheta}{\tau}\Reward(\tau)}
        %\int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
\end{align}
Notice that the distribution defining the gradient is induced by the current policy. This aspect introduces a nonstationarity in the sampling process. Since the underlying distribution changes over time, it is necessary to resample at each update or use weighting techniques such as importance sampling.
Here, we consider the \emph{online learning scenario}, where trajectories are sampled by interacting with the environment at each policy change. 
In this setting, stochastic gradient ascent is typically employed.
At each iteration $k >0$, a batch $\mathcal{D}_N^k = \{\tau_i\}_{i=0}^N$ of $N>0$ trajectories is collected using policy $\pi_{\vtheta_k}$.
The policy is then updated as $\vtheta_{k+1}  = \vtheta_k + \alpha\gradApp{\vtheta_k}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of~\eqref{E:policygradient} using $\mathcal{D}_N^k$, \ie
\begin{align} \label{E:policygradient.estimate}
        \gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N} g(\tau_i|\vtheta), \quad \tau_i \in \mathcal{D}_N^k.
\end{align}
The most widespread gradient estimators are REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
For example, REINFORCE is such that $g(\tau|\vtheta) = \nabla \log\pi(\tau|\vtheta) \mathcal{R}(\tau)$. 
Although REINFORCE definition is simpler, G(PO)MDP is often preferred due to its lower variance.
We refer the reader to App.~\ref{A:gradient_estimators} for details.

@@

The idea of SVRG is to alternate full and stochastic gradient updates. 
Each $m = \mathcal{O}(N)$ iterations, a snapshot $\widetilde{\vtheta}$ of the current parameter is saved together with its full gradient $\nabla f(\widetilde{\vtheta}) = \frac{1}{N} \sum_i z(x_i|\widetilde{\vtheta})$.
Between snapshots, the parameter is updated with a gradient estimate corrected using stochastic gradient, for any $t \in \{0,\ldots,m-1\}$
\begin{equation}\label{E:svrg.gradient.correction}
        \blacktriangledown f(\vtheta_{t}) = \nabla f(\tilde{\vtheta}) + \nabla z(x_i | \vtheta_t) - \nabla z(x_i | \tilde{\vtheta}),
\end{equation} 
where $x_i$ is sampled uniformly at random from $\mathcal{D}_N$ (\ie $x_i \sim \mathcal{U}(\mathcal{D}_N)$).
Note that $t=0$ corresponds to a FG step (\ie $\blacktriangledown f(\vtheta_0) = \nabla f(\wt{\vtheta})$) since $\vtheta_0 := \wt{\vtheta}$.
The corrected gradient $\blacktriangledown f(\vtheta)$ is an unbiased estimate of $\nabla f(\vtheta)$, and it is able to control the variance introduced by sampling even with a fixed step size, without resorting to plain full gradient. The resulting IFO complexity is $\mathcal{O}(N+\nicefrac{\sqrt{N}}{\epsilon})$ \cite{reddi2016stochastic}.

More recently, some extensions of variance reduction algorithms to the non-convex objectives have been proposed~\citep[\eg][]{reddi2016stochastic,allen2016variance,reddi2016stochastic,reddi2016fast}. In this scenario, $f$ is typically required to be L-smooth, \ie $\norm[]{\nabla f(\vtheta') - \nabla f(\vtheta)} \leq L\norm[]{\vtheta'-\vtheta}$ for each $\vtheta,\vtheta'\in\realspace^n$ and for some Lipschitz constant $L$. Under this hypothesis, the IFO complexity of full gradient  and of stochastic gradient ascent are the same as in the \textcolor{red}{concave} case~\citep{nesterov2013introductory,ghadimi2013stochastic}. Also in this case, SVRG yields an advantageous trade-off, with an IFO complexity of $\mathcal{O}(N + \nicefrac{N^{\nicefrac{2}{3}}}{N})$~\citep{reddi2016stochastic}. The only additional requirement is to select $\vtheta^*$ uniformly at random among all the $\vtheta_k$ instead of simply setting it to the final value. Pseudocode of SVRG for both the convex and the non-convex case is provided in Alg.~\ref{alg:svrg}.

\begin{algorithm}[tb]
	\caption{SVRG}
	\label{alg:svrg}
	\begin{algorithmic}
            \STATE {\bfseries Input:} a dataset $\mathcal{D}_N$, number of epochs $S$, epoch size $m$, step size $\alpha$, initial parameter $\vtheta_{0}^0:=\wt{\vtheta}^0$
		\FOR{$s=0$ {\bfseries to} $S-1$}
        \STATE $\wt{\mu} = f(\wt{\vtheta}^s)$
		\FOR{$t=0$ {\bfseries to} $m-1$}
%		\STATE Pick $i$ uniformly at random from $[1,N]$
        \STATE $x_i \sim \mathcal{U}\left(\mathcal{D}_N\right)$
		\STATE $v^{s}_t = 
%			\nabla f(\wt{\vtheta}^s) + 
			\wt{\mu} + 
			\nabla g(x_i|\vtheta_t^{s}) -
			\nabla g(x_i|\wt{\vtheta}^{s})
		$
        \STATE $\vtheta_{t+1}^{s} = \vtheta_t^{s} + \alpha v^{s}_t$
		\ENDFOR
        \STATE $\vtheta_0^{s+1} := \wt{\vtheta}^{s+1} = \vtheta_{m}^s$
		\ENDFOR
        \STATE \underline{Concave case:} \textbf{return} $\vtheta_{0}^S$
        \STATE \underline{Non-Concave case:} \textbf{return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $\{[0,S[\times[0,m[\} \cup \{(S,0)\}$
	\end{algorithmic}
\end{algorithm}

@@

We can now introduce Stochastic Variance-Reduced Policy Gradient (SVRPG) for a generic policy gradient estimator $g$ (see Eq.~\ref{E:policygradient.estimate}).
% For simplicity, we focus on the REINFORCE estimator but the extension to G(PO)MDP is straightforward.
Between snapshots, the SVRPG estimate is corrected using importance weighting, for any $t \in \{0,\ldots,m-1\}$
\begin{align*}
        \blacktriangledown J(\vtheta_{t}) &= \wh{\nabla}_N J(\wt{\vtheta}) + g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
\end{align*}
% \begin{align*}
% 	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) \\
% 		&+ \nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
% 		- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i),
% \end{align*}
where $\wh{\nabla}_N J(\wt{\vtheta})$ is as in~\eqref{E:policygradient.estimate} where $\mathcal{D}_N$ is sampled using the snapshot policy $\pi_{\wt{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pi_{\vtheta_t}$, and $\omega(\tau|\vtheta_t, \wt{\vtheta}) = \frac{\pi(\tau|\wt{\vtheta})}{\pi(\tau|\vtheta_t)}$ is an importance weight from $\pi_{\vtheta_t}$ to the snapshot policy $\pi_{\wt{\vtheta}}$. 
Similarly to SVRG, we have that $\vtheta_0 := \wt{\vtheta}$ and the update is a FG step.
Our update is still fundamentally on-policy, since the weighting concerns only the correction term. However, this partial ``off-policyness'' is enough to introduce variance. This is a well-known issue of importance sampling~\citep[\eg][]{thomas2015high}. To mitigate it, we use mini-batches of trajectories of size $B \ll N$ to average the correction, \ie
\begin{align}\label{E:svrpg.estimate.batch}
        \blacktriangledown J(\vtheta_{t}) &:= v_t= \wh{\nabla}_N J(\wt{\vtheta})\\ \notag
                                            & \quad{} + \frac{1}{B} \sum_{i=0}^{B-1} \left[
        g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
        \right].
\end{align}
%\begin{align*}
%&\blacktriangledown J(\vtheta) = \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) +\\
%&\quad\frac{1}{B}\left[\sum_{i=0}^{B-1}
%\nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
%- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i)\right].
%\end{align*}
The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case~\citep{harikandeh2015stopwasting,konevcny2016mini}. It is easy to show that the SVRPG estimator has the following, desirable properties (see App.~\ref{app:proofs}):
\begin{lemma}
Denote by $\wh{\nabla}_N^{\textsc{RF}} J(\vtheta)$ the REINFORCE estimator in~\eqref{E:policygradient.estimate} and let $\vtheta^* \in \argmin_{\vtheta} \{J(\vtheta)\}$. Then, the SVRG estimate in~\eqref{E:svrpg.estimate.batch} is \emph{unbiased}
\[
\mathop{\mathbb{E}}
\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
\]
and regardless of the mini-batch size $B$:
\[
	\mathbb{V}ar\left[\gradBlack{\vtheta^*}\right] = 
    \mathbb{V}ar\left[\wh{\nabla}_N^{\textsc{RF}}J(\vtheta^*)\right].
\]
\todopir{Definition of Var??}
\end{lemma}

%\begin{restatable}{lemma}{unbias}\label{lemma:unbias}
%The SVRPG estimator is unbiased:
%\[
%\mathop{\mathbb{E}}
%\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
%\]
%\end{restatable}
%\begin{restatable}{lemma}{zerovar}\label{lemma:zerovar}
%At convergence, the variance of the SVRPG estimator is equal to the variance of the REINFORCE estimator, regardless of the mini-batch size $B$:
%\[
%	\mathbb{V}ar\left[\gradBlack{\vtheta^*}\right] = 
%	\mathbb{V}ar\left[\gradApp{\vtheta^*}{N}\right].
%\]
%\end{restatable}
In particular, the latter result suggests that an SVRG-like algorithm using $\gradBlack{\vtheta}$ can actually achieve faster convergence, by performing much more parameter updates with the same data without introducing additional variance (at least asymptotically).
The resulting SVRPG policy optimization method is detailed in Alg.~\ref{alg:svrpg}.
To obtain a method that can be used in practice, a number of additional details must be specified, namely the choice of epoch size $m$, the number $N$ of trajectories used to estimate the full gradient, the mini-batch size $B$, and the step size $\alpha$.
Adaptive selection of these meta-parameters is also possible, and is explored in Sec.~\ref{sec:stopping}.
However, the next section will provide general converge guarantees for SVRPG with REINFORCE estimator.

\begin{algorithm}[tb]
        \caption{SVRPG}
	\label{alg:svrpg}
	\begin{algorithmic}
        \STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, batch size $N$, mini-batch size $B$, gradient estimator $g$, initial parameter $\vtheta_{0}^0 := \wt{\vtheta}^0$
		\FOR{$s=0$ {\bfseries to} $S-1$}
        \STATE Sample $N$ trajectories $\{\tau_j\}$ from $\pi(\cdot\vert\tilde{\vtheta}^{s})$
        \STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\ref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
		\FOR{$t=0$ {\bfseries to} $m-1$}
        \STATE Sample $B$ trajectories $\{\tau_i\}$ from $\pi(\cdot\vert\vtheta_t^{s+1})$
        \STATE $c^s_t = \frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^s) - \omega(\tau_i|\vtheta^s_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta})^s\right)$
%        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1}
%        \begin{aligned}[t]
%                & \Reward(\tau_i) \Big[ \score{}{\tau_i|\vtheta_t^{s+1}} \\
%                &- \omega(\tau|\vtheta^{s+1}_t, \wt{\vtheta}^s)\score{}{\tau_i|\wt{\vtheta}^{s}} \Big]
%        \end{aligned}$
        \STATE $v^{s}_t = \wt{\mu} + c^s_t$ % \gradApp{\wt{\vtheta}^{s}}{N} + c^{s}_t$
%%		\begin{align*}
%%		\blacktriangledown J(\vtheta_t^{s+1}) = 
%%		&\gradApp{\tilde{\vtheta}^s}{N} \\
%%		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
%%		\score{}{\tau_i\vert\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
%%		&\left. - \omega(\tau_i)\score{}{\tau_i \vert \tilde{\vtheta}^{s}}\Reward(\tau_i)\right]
%%		\end{align*}
        \STATE $\vtheta_{t+1}^{s} = \vtheta_t^{s} + \alpha v^{s}_t$
		\ENDFOR
        \STATE $\vtheta_0^{s+1} := \tilde{\vtheta}^{s+1} = \vtheta_{m}^s$
		\ENDFOR
        \STATE {\bfseries return} $\vtheta_t^s$ with $(s,t)$ picked uniformly at random from $\{[0,S[\times[0,m[\} \cup \{(S,0)\}$
	\end{algorithmic}
\end{algorithm}

\section{Convergence Guarantees of SVRPG}\label{sec:conv}
In this section, we state the convergence guarantees for SVRPG with REINFORCE estimator (refer to Alg.~\ref{alg:svrpg} with $g(\tau|\vtheta) =  \nabla \log\pi(\tau|\vtheta) \mathcal{R}(\tau)$).\footnote{Similar results can be derived for G(PO)MDP, see App.~\ref{app:}.}
We mainly leverage on the recent analysis of \textcolor{red}{non-concave} SVRG~\cite{reddi2016stochastic,allen2016variance}.
Each of the three challenges presented at the beginning of Sec.~\ref{sec:alg} can potentially prevent convergence, so we need additional assumptions.
In App.~\ref{app:gauss} we show how the widespread Gaussian policy satisfies these assumptions.

@@

We can now state the convergence guarantees for SVRPG.
\begin{restatable}[Convergence of the REINFORCE-SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
Consider the REINFORCE estimator in Eq.~\ref{E:policygradient.estimate}.
Under Ass.~\ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Alg.~\ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
\begin{align*}
	&\EVV[]
	{\norm[2]{\nabla J(\vtheta_A)}^2} 
		\leq
		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
		\frac{\zeta}{N}
		+\frac{\xi}{B}
\end{align*}
\end{restatable}
% Specific constants and meta-parameter constraints are provided in the proof of the theorem in Appendix \ref{app:proofs}.
Refer to App.~\ref{app:proofs} for a detailed proof involving definition of the constants and the meta-parameter constraints.
%The theoretical settings may be too conservative.
In Sec.~\ref{sec:stopping} we propose a joint selection of step size $\alpha$ and epoch size $m$ which is more suited for applications. 
By analysing the upper-bound in Thm.~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on \textcolor{red}{non-concave} SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
To achieve true convergence, both the batch size $N$ and the mini-batch size $B$ should increase over time. 
In practice, it is enough to choose $N$ and $B$ large enough to make the second and the third term negligible, \ie to mitigate the variance introduced by FG approximation and importance sampling, respectively.
Since the optimal configuration may be task-dependent, we leave this choice to the empirical analysis of Sec.~\ref{sec:exp}.

\textbf{G(PO)MDP.} A similar result to the one stated in Thm~\ref{theo:convergence} can be derived for the G(PO)MDP estimator. A sketch of the proof is reported in~\ref{app:gpomdpproof}.
In practice, due to the lower variance G(PO)MDP is preferred to REINFORCE.
For this reason we will use G(PO)MDP for the experiments (see Sec.~\ref{sec:exp}).
\todopir{sketch of proof for GPOMDP in appendix}
